Near Optimal Behavior with Approximate State Abstraction

### 2016-02-03 paper/theory_results.tex
Starting to unify notation

% Intro to approx state abstraction results.
Here, we introduce our formal method for analyzing approximate state abstraction, including results bounding the error associated with these abstraction methods. In particular, we demonstrate that abstractions based on $Q^*$ similarity, model similarity (that is, $\mathcal{T}$ and $\mathcal{R}$), and similarity between distributions over $Q^*$, for both multinomial and Boltzmann distributions, induce abstract \acp{MDP} for which the optimal policy has bounded error in the ground MDP.

In general, our proof strategy for each of the included results is as follows:
\begin{itemize}
\item First, we relate the values of states in the ground MDP to the value of states in the abstract \ac{MDP},
\item then, we relate the value of states in the ground \ac{MDP} to the value of states in the abstract \ac{MDP} under the optimal abstract policy.
\end{itemize}

% Explanation of the above.
The first statement is relating Q values defined over the ground MDP with Q values defined over the abstract MDP, while the second statement is comparing the optimal Q values in the ground MDP, with the Q values in the ground MDP under the abstract policy. This is because our learners will solve for the optimal policy in the abstract, and use it in the ground MDP by mapping each ground state to its abstract state and querying the abstract policy for behavior. That is:
\begin{equation}
\pi_G(s) = \pi_A(\phi(s))
\end{equation}

@@

We now introduce some additional notation to formally discuss these results.

{\bf $\ep_f$ definition:} Given a function $f: \mathcal{S}_G \times \mathcal{A} \rightarrow \mathcal{R}$ and a fixed non-negative $\epsilon \in \mathbb{R}$, we define $\ep_f$ as a type of approximate state-aggregation function that satisfies the following for any two ground states $s_1$ and $s_2$: 
\begin{equation}
\label{eq:phi_f}
\ep_f(s_1) = \ep_f(s_2) \rightarrow \forall_a \left|f(s_1, a) - f(s_2, a)\right| \leq \epsilon
\end{equation}

That is, when $\ep_f$ aggregates states, all aggregated states have values of $f$ within $\epsilon$ of each other for all actions. Note that, for any $f$, there are many functions that satisfy this criterion.

@@

% Theorem 1
\begin{thm}
There exist at least four classes of approximate state-aggregation functions, $\ep_{Q^*}$, $\ep_{\text{model}}$, $\ep_{\text{bolt}}$ and $\ep_{\text{mult}}$ for which the optimal policy in the abstract \ac{MDP}, applied to the ground \ac{MDP}, has suboptimality bounded polynomially in $\epsilon$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq poly(\epsilon).
\end{equation}

\end{thm}

@@

% Subsection: Optimal Q Function
\subsection{Optimal Q Function: $\ep_{Q^*}$}

@@

%Claim 2
\begin{clm}
\label{clm:optAbsActionNearOptGround}
The optimal action in the abstract MDP has a Q value in the ground which is nearly optimal.
\end{clm}

@@

% Subsection: Model Similarity
\subsection{Model Similarity: $\ep_{model}$}

@@

% Subsection: Multinomial over Optimal Q
\enote{If not here, we should have some motivation for multinomial (and boltz)}
\subsection{Multinomial over Optimal Q: $\ep_{mult}$}

@@

% Subsection: Boltzmann over Optimal Q
\subsection{Boltzmann over Optimal Q: $\ep_{bolt}$}

{\bf Lemma 4:} When $S_A$ is created using a function of the $\ep_{bolt}$ type, for some constant positive $k \in \mathbb{R}$::
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{\epsilon(\textsc{QMax}|\mathcal{A}| + k\epsilon) + A}{1-\gamma}:
\end{equation}

% There are at least four classes of functions:
% 1 \phi_q^*
% 2 \phi_model
% 3 \phi_mult
% 4 \phi_bolt

% Explain each of the Phis

% One theorem, each separate phi is its own lemma.

% Proof of Q^*

>>>>

% Intro to approx state abstraction results.
Here, we introduce our formal method for analyzing approximate state abstraction, including results bounding the error associated with these abstraction methods. In particular, we demonstrate that abstractions based on $Q^*$ similarity (\ref{sec:Q*}), model similarity (\ref{sec:model}), and similarity between distributions over $Q^*$, for both Boltzmann (\ref{sec:boltz}) and multinomial (\ref{sec:mult}) distributions, induce abstract \acp{MDP} for which the optimal policy has bounded error in the ground MDP.

%In general, our proof strategy for each of the included results is as follows:
%\begin{itemize}
%\item First, we relate the values of states in the ground MDP to the value of states in the abstract \ac{MDP},
%\item then, we relate the value of states in the ground \ac{MDP} to the value of states in the abstract \ac{MDP} under the optimal abstract policy.
%\end{itemize}

% Explanation of the above.
%The first statement is relating Q values defined over the ground MDP with Q values defined over the abstract MDP, while the second statement is comparing the optimal Q values in the ground MDP, with the Q values in the ground MDP under the abstract policy. This is because our learners will solve for the optimal policy in the abstract, and use it in the ground MDP by mapping each ground state to its abstract state and querying the abstract policy for behavior. That is:
%\begin{equation}
%\pi_G(s) = \pi_A(\phi(s))
%\end{equation}

@@

We now introduce some additional notation to formally discuss these results:

% Abstract policy in ground.
We are interested in how the optimal policy in the abstract \ac{MDP} performs in the ground \ac{MDP}. To evaluate a policy $\pi_A:\mathcal{A}\rightarrow\mathcal{A}$, we map a ground state into the abstract state space, and query the abstract policy for behavior. Specifically:
\begin{equation}
V^{\pi_A}(s) = \mathcal{R}_G(s,\pi_A(\phi(s))) + \gamma \sum_{s' \in \mathcal{S}_G} \mathcal{T}_G(s,\pi_A(\phi(s)),s')V^{\pi_A}(s').
\end{equation}
\dnote{I don't think this is right. Should be the max action? Or am I confused? My reading of the Q function from above is then possibly confused...}

\begin{defn}
Given a function $f: \mathcal{S}_G \times \mathcal{A} \rightarrow \mathcal{R}$ and a fixed non-negative $\epsilon \in \mathbb{R}$, we define $\ep_f$ as a type of approximate state-aggregation function that satisfies the following for any two ground states $s_1$ and $s_2$: 
\begin{equation}
\label{eq:phi_f}
\ep_f(s_1) = \ep_f(s_2) \rightarrow \forall_a \left|f(s_1, a) - f(s_2, a)\right| \leq \epsilon
\end{equation}
\end{defn}

That is, when $\ep_f$ aggregates states, all aggregated states have values of $f$ within $\epsilon$ of each other for all actions. Note that, for any $f$, there are many functions that satisfy the $\ep_f$ criterion.

@@

% Theorem 1
\begin{thm}
There exist at least four classes of approximate state-aggregation functions, $\ep_{Q^*}$, $\ep_{\text{model}}$, $\ep_{\text{bolt}}$ and $\ep_{\text{mult}}$ for which the optimal policy in the abstract \ac{MDP}, applied to the ground \ac{MDP}, has suboptimality bounded polynomially in $\epsilon$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq poly(\epsilon).
\end{equation}
\end{thm}

@@

% Subsection: Optimal Q Function
\subsection{Optimal Q Function: $\ep_{Q^*}$}
\label{sec:Q*}

@@

%Claim 2
\begin{clm}
\label{clm:optAbsActionNearOptGround}
The optimal action in the abstract MDP has a Q value in the ground which is nearly optimal:
\end{clm}

@@

% Subsection: Model Similarity
\subsection{Model Similarity: $\ep_{model}$}
\label{sec:model}

@@

% Subsection: Boltzmann over Optimal Q
\enote{Motivate Boltzmann + multinomial}

\subsection{Boltzmann over Optimal Q: $\ep_{bolt}$}
\label{sec:boltz}

{\bf Lemma 4:} When $S_A$ is created using a function of the $\ep_{bolt}$ type, for some constant positive $k \in \mathbb{R}$::
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq 
\epsilon:
\end{equation}

% Subsection: Multinomial over Optimal Q
\subsection{Multinomial over Optimal Q: $\ep_{mult}$}
\label{sec:mult}

@@
